## Tasks on the R datatypes

------------------------------------------------------------------------

### Datatypes

**There are 5 datatypes:**

```{r}
# numeric (decimal, double precision)
numeric1 <- c(0.5, -0.2)
numeric1
# logical (both forms acceptable)
logical1 <- c(T, FALSE)
logical1
# character
character1 <- c("ABC", "d", "EF", "#$&^*")
character1
# integer
integer1 <- c(0:13)
integer1
# complex
complex1 <- c(1 + 2i, 3 + 10i)
complex1
```

> it can be converted to the specified type

```{r}
class(numeric1)
numeric1 <- as.character(numeric1)
numeric1
class(numeric1)
```

> To get an integer type like **3L**

```{r}
x <- 3L
x
class(x)
```

------------------------------------------------------------------------

### Lists and vectors

#### *Vectors (Only one datatype)*

> If there is more than one datatype in vector, coersion occurs. It may cause \> error or not.

```{r}
# function c() creates vector
vector1 <- c(1.2, 1.3, 1.4)
vector1
```

> **":"** can be used for **integer** sequence only!

```{r}
# ":" can be used for integer sequence only! 
vector2 <- c(1:13)
vector2
```

The other way to create vector (creates empty vector):

```{r}
vector3 <- vector("character", length = 15)
vector3
```

#### *Lists (different datatypes within one list, list is also considered as vector)*

> Lists have elements' indexing from 1

```{r}
list1 <- list(1:5, "Ef", "$#", 1.2, 1 + 4i)
list1
```

```{r}
list1[3]
```

------------------------------------------------------------------------

### Matrices (special type of *vector*)

Empty matrix:

```{r}
matr1 <- matrix(nrow = 4, ncol = 4)
matr1

#matrix dimensionality
dim(matr1)

#attributes, dimensionality is first
attributes(matr1)

```

Creating filled-in matrix

```{r}
matr2 <- matrix(-1:14, nrow = 4, ncol = 4)
matr2
attributes(matr2)
```

> Matrix can also be created from vector

```{r}
values <- -1:14
dim(values) <- c(4, 4)
values
```

Construct matrix by **binding**:

> **Column-binding:**

```{r}
col1 <- 1:10
col2 <- 11:20
col3 <- 21:30
cbind(col1, col2, col3)
```

> **Row-binding**

```{r}
row1 <- 1:10
row2 <- 11:20
row3 <- 21:30
rbind(col1, col2, col3)
```

------------------------------------------------------------------------

### Factors (to represent *categorical* data)

Vector of integers, where each *integer* has a *label* (categorical data).

```{r}
factor1 <- factor(c('category #1', 'category #2',  'category #1',  'category #3'))
factor1
# number of occurencies of each factor
table(factor1)
# convert to integer representation (change attribute level)
unclass(factor1)
```

> When you want to specify baseline, you can specify the first level

```{r}
factor2 <- factor(c('category #1', 'category #2',  'category #1',  'category #3'),
                  levels = c('category #1', 'category #2','category #3'))
factor2
```

### Missing values (*NA* and *NaN*)

Determine if objects are **NA**:

```{r}
# take previously created empty (NA) matrix
is.na(matr1)
```

Determine if objects are **NaN**:

```{r}
# the same matrix
is.nan(matr1)
```

> **NA** != **NaN**


### Data Frames (tabular data)


